---
title: "Neurodevelopmental Disorders Genes"

author:
- name: Author Nicolò Caporale (nicolo.caporale@gmail.com)
- name: Note the multintersect function was written by Pierre-Luc Germain (pl.germain@gmail.com)

date: 'Date: `r format(Sys.Date(), "%B %d, %Y")`'
output:
    html_document:
        toc: true
        toc_float: true
        theme: 'simplex'
        highlight: 'espresso'
        #code_folding: hide
---

If you have 1 or more lists of genes and you want to know how relevant they are in the context of neurodevelopment and autism, here you can find a very simple pipeline that can help you.

This is a very simple analysis but it is **standing on the shoulders of giants**, as you can see from the references. Most of this work, indeed, is based on literature search, past experience with public dataset mining, manual selection and curation of the most relevant studies dealing with the genetics of neurodevelopmental disorders. It is by definition not comprehensive and it will need constant curation to be kept updated: so I ask you to contact me for constructive comments and suggestion to improve it..


### Data loading

```{r, warning=FALSE, message=FALSE}
source("Functions/Functions.R")
load("Data/Nowakowski_et_al.RData")
load("Data/Organoids.RData")
load("Data/universe.RData")
load("Data/ASD.RData")
```


## Overlaps with ASD genes

I take 2 gene lists of my interest and intersect them with all the relevant ASD genes. As an example I take here the genes that define cell populations of different neural differentiation stages, from a single cell analysis of cortical brain organoids that we did in our lab. 
https://www.biorxiv.org/content/10.1101/484741v2

```{r,warning=FALSE,message=FALSE,fig.height=7,fig.width=8}
multintersect(ClusterInternalArea, ASD, universe = universe)

```

## Expression of ASD relevant genes

As an example I plot ASD relevant genes across organoid samples that were exposed to different treatments
```{r,warning=FALSE,message=FALSE,fig.height=7,fig.width=8}
load("Data/treatment.RData")

byheatmap
```

### Data download and praparation

For details on data preparation look [here](./DataPreparation/ASD/DataPreparation.html) for ASD genes

***
[here](./DataPreparation/Nowakowski_et_al/DataPreparation.html) for fetal brain single cell clusters

***
[here](./DataPreparation/Lopez_et_al/DataPreparation.html) for cortical brain organoids single cell clusters
