---
date: 'Date: `r format(Sys.Date(), "%B %d, %Y")`'

output:
    html_document:
        #toc: true
        #toc_float: true
        theme: 'simplex'
        highlight: 'espresso'
        #code_folding: hide
---


To go back to the main page go [here](../../NDD.html)


## Data preparation
SFARI genes were downloaded from https://gene.sfari.org/database/gene-scoring/

```{r}
SFARI <- read.csv("SFARI-Gene_genes_06-20-2019release_07-13-2019export.csv")
SFARIgenes <- list()
SFARIgenes$score1 <- as.character(SFARI$gene.symbol[which(SFARI$gene.score==1)])
SFARIgenes$score2 <- as.character(SFARI$gene.symbol[which(SFARI$gene.score==2)])
SFARIgenes$score3 <- as.character(SFARI$gene.symbol[which(SFARI$gene.score==3)])
SFARIgenes$score4 <- as.character(SFARI$gene.symbol[which(SFARI$gene.score==4)])
SFARIgenes$score5 <- as.character(SFARI$gene.symbol[which(SFARI$gene.score==5)])
SFARIgenes$score6 <- as.character(SFARI$gene.symbol[which(SFARI$gene.score==6)])
SFARIgenesAll<- unlist(SFARIgenes)
```

AutismKB genes were downloaded from http://db.cbi.pku.edu.cn/autismkb_v2/index.php
```{r}
AutismKB <- read.table("AutismKB.txt", header = F, sep = "\t")
colnames(AutismKB) <- c("Gene Symbol",	"Entrez ID",	"Genome-Wide Association Studies",	"CNV/SV Studies",	"Linkage Analyses",	"Low-Scale Genetic Association Studies",	"Expression Profilings",	"NGS de novo Mutation Studies",	"NGS Mosaic Mutation Studies",	"NGS Other Studies",	"Low-Scale Gene Studies",	"total score")
AutismKBgenes<- as.character(AutismKB$`Gene Symbol`)
```

DDD genes were downloaded from https://decipher.sanger.ac.uk/about#downloads/data
```{r}
DDD <- read.csv("DDG2P_16_7_2019.csv")
DDDneuro <- DDD[c(grep("autism", DDD$disease.name,ignore.case = T),grep("intellectual", DDD$disease.name,ignore.case = T),grep("neurodevelopm", DDD$disease.name, ignore.case = T)),]
DDDneurogenes <- as.character(DDDneuro$gene.symbol)
```

MSSNG genes were downloaded from https://www.ncbi.nlm.nih.gov/pubmed/28263302 Supplementary Table 5
```{r}
library(readxl)
MSSNG <- suppressMessages(read_excel("nn.4524-S7.xlsx"))
colnames(MSSNG) <- MSSNG[1,]
MSSNG <- MSSNG[-1,] 
MSSNGgenes <- as.character(MSSNG$gene_symbol)
```

ASDsevere genes were downloaded from https://www.nature.com/articles/s41436-018-0380-2 Supplementary Table S6
```{r}
ASDsevere <- suppressMessages(read_excel("41436_2018_380_MOESM7_ESM.xlsx"))
ASDsevereGenes <- as.character(ASDsevere$Gene)
```

ID genes were downloaded from https://www.nature.com/articles/ng.3792 Supplementary Table S2
```{r}
ID <- suppressMessages(read_excel("41588_2017_BFng3792_MOESM20_ESM.xlsx"))
colnames(ID) <- ID[1,]
ID <- ID[-1,] 
IDGenes <- as.character(ID$Gene)
```

OMIM genes were downloaded from https://www.omim.org/downloads/
I then select only the genes that are related to neurodevelopmental phenotypes
```{r}
morbidmap <- read.table("morbidmap.txt",sep='\t', comment.char='#',fill = T, header = F,quote = "" )
colnames(morbidmap) <-c("Phenotype",	"Gene Symbols",	"MIM Number",	"Cyto Location")

#run this chunck if you want to check the specific phenopyes selected
#morbidmap$Phenotype[c(grep("autism", morbidmap$Phenotype,ignore.case = T),grep("intellectual", morbidmap$Phenotype,ignore.case = T),grep("neurodevelopm", morbidmap$Phenotype,ignore.case = T))]

NeuroOmim <- morbidmap[c(grep("autism", morbidmap$Phenotype,ignore.case = T),grep("intellectual", morbidmap$Phenotype,ignore.case = T),grep("neurodevelopm", morbidmap$Phenotype,ignore.case = T)),]

NeuroOmimGenes <- as.character(morbidmap$`Gene Symbols`[c(grep("autism", morbidmap$Phenotype,ignore.case = T),grep("intellectual", morbidmap$Phenotype,ignore.case = T),grep("neurodevelopm", morbidmap$Phenotype,ignore.case = T))])
NeuroOmimGenes <- strsplit(NeuroOmimGenes,split = ',')
NeuroOmimGenes <- unique(unlist(NeuroOmimGenes))
```

GDI indexes were downloaded from http://lab.rockefeller.edu/casanova/GDI
```{r}
GDI <- read.table("GDI_full_10282015.txt", header = T, sep = "\t")
```

I then save all the objects
```{r}
ASD <- list(SFARI=SFARIgenesAll, AutismKB=AutismKBgenes, DDD=DDDneurogenes, MSSNG=MSSNGgenes, ASDsevere=ASDsevereGenes, ID=IDGenes, OMIM= NeuroOmimGenes)
save(ASD, SFARI, SFARIgenes, AutismKB, DDDneuro, MSSNG, ASDsevere, ID, NeuroOmim, GDI, file = "../../Data/ASD.RData")
```

